---
title: "roahd 1.4.3"
description: |
  An R package for the robust analysis of high-dimensional data.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2022-01-31
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - functional data
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="50%", fig.align='center'}
img <- png::readPNG("../../images/roahd_hex_trans.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

## Overview

The [**roahd**](https://astamm.github.io/roahd/index.html) (Robust Analysis of
High-dimensional Data) package [@Ieva2019] allows to use a set of statistical
tools for the exploration and robustification of univariate and multivariate
functional data sets through the use of depth-based statistical methods.

In the implementation of functions, special attention was put to their
efficiency, so that they can be profitably used also for the analysis of
high-dimensional datasets.

For a full-featured description of the package, please take a look at the
[**roahd**](https://astamm.github.io/roahd/articles/roahd.html) vignette.

## New feature

We added tools for manipulating and visualizing depthgrams
[@aleman2021visualizing]. This mathematical constructs aim at facilitating the
visualization of outliers in high dimensional functional data sets. The
[`depthgram()`](https://astamm.github.io/roahd/reference/depthGram.html)
function computes a number of depthgrams from the functional data set. An S3
specialized method for
[`plot()`](https://astamm.github.io/roahd/reference/plot.depthgram.html) makes
it possible to visualize the depthgrams and proceed with a visual inspection at
outliers.
