---
title: "New paper on walking activity recognition"
description: |
  Check out our newly published research which presents a unified framework to 
  perform walking activity recognition from rotation time series collected by a 
  wearable sensor in a free-living environment.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2022-05-09
output:
  distill::distill_article:
    self_contained: false
categories:
  - research
  - rotation-valued functional data
  - machine learning
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="80%", fig.align='center'}
img <- png::readPNG("../../images/sensors_warm.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

Solutions to assess walking deficiencies are widespread and largely used in
healthcare. Wearable sensors are particularly appealing, as they offer the
possibility to monitor gait in everyday life, outside a facility in which the
context of evaluation biases the measure. While some wearable sensors are
powerful enough to integrate complex walking activity recognition models,
non-invasive lightweight sensors do not always have the computing or memory
capacity to run them. In this paper, we propose a walking activity recognition
model that offers a viable solution to this problem for any wearable sensors
that measure rotational motion of body parts. Specifically, the model was
trained and tuned using data collected by a motion sensor in the form of a unit
quaternion time series recording the hip rotation over time.

This time series was then transformed into a real-valued time series of geodesic
distances between consecutive quaternions. Moving average and moving standard
deviation versions of this time series were fed to standard machine learning
classification algorithms. To compare the different models, we used metrics to
assess classification performance (precision and accuracy) while maintaining the
detection prevalence at the level of the prevalence of walking activities in the
data, as well as metrics to assess change point detection capability and
computation time. Our results suggest that the walking activity recognition
model with a decision tree classifier yields the best compromise in terms of
precision and computation time.

The sensor that was used had purposely low computing and memory capacity so that
reported performances can be thought of as the lower bounds of what can be
achieved. Walking activity recognition is performed online, i.e., on-the-fly,
which further extends the range of applicability of our model to sensors with
very low memory capacity.

Find out more by reading our paper [@brard2022novel] which can be found at
<https://doi.org/10.3390/s22093555>.
