---
title: "riot 0.0.1"
description: |
  An R interface for popular tractography file formats.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2022-05-09
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - functional data
  - connectome data
  - data import
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="50%", fig.align='center'}
img <- png::readPNG("../../images/riot_hex_trans.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

## Overview

The [**riot**](https://astamm.github.io/riot/index.html) (R Input/Output for
Tractography) package provides an R interface for importing and exporting
tractography data to and from `R`. Currently supported importing formats are:

- native [VTK](https://vtk.org) `.vtk` and `.vtp` files; and,
- [medInria](https://med.inria.fr) `.fds' files.

The package reads tractography data into a
[tibble](https://tibble.tidyverse.org) in which each row is a point
characterized by at least the following five variables:

- `X`, `Y`, `Z`: 3D coordinates of the current point;
- `PointId`: Identification number of the current point among all points of the
streamline it belongs to;
- `StreamlineId`: Identification number of the streamline which the current
point belongs to.

The points might also have attributes or a color assigned to them, in which
case, additional variables will be properly created to import them as well. The
user can perform statistical analysis on the point cloud and store any new
variable that (s)he would deem to be useful as additional column of the
[tibble](https://tibble.tidyverse.org). The packages also allow to write back
the [tibble](https://tibble.tidyverse.org), including all newly created
attributes, into the following exporting formats:

- native [VTK](https://vtk.org) `.vtk` and `.vtp` files; or,
- [medInria](https://med.inria.fr) `.fds` files.

## Installation

You can install the package directly from CRAN:
```r
install.packages("riot")
```
or you can choose to install the development version from GitHub:
```r
# install.packages("remotes")
remotes::install_github("astamm/riot")
```

The package has its own webpage: https://astamm.github.io/riot/.
