---
title: "riot 0.0.1"
description: |
  An R interface for Popular Tractography File Formats
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2022-05-09
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - connectomics
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="50%", fig.align='center'}
img <- png::readPNG("../../images/riot_hex_trans.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

## Overview

The [`riot`](https://astamm.github.io/riot/index.html) (R Input/Output for Tractography) package provides an R interface to import tractography data in R. Currently supported formats for importing are native [VTK](https://vtk.org) `.vtk` and `.vtp` files as well as [MRtrix](https://mrtrix.readthedocs.io/en/latest/getting_started/image_data.html) `.tck/.tsf` files and [TrackVis](http://trackvis.org/docs/?subsect=fileformat) `.trk` files.

## Installation

You can install the package directly from CRAN:
```
install.packages("riot")
```
or you can choose to install the development version from GitHub:
```
# install.packages("remotes")
remotes::install_github("astamm/riot")
```

The package has its own webpage: https://astamm.github.io/riot/.
