---
title: "squat 0.1.0"
description: |
  An implementation of statistical tools for the analysis of unit quaternion 
  time series. It relies on pre-existing quaternion data structure provided by 
  the 'Eigen' C++ library.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2023-01-03
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - rotation-valued functional data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r logo, out.width="80%", fig.align='center'}
img <- magick::image_read("../../images/squat_anim.gif")
img
```


## Overview

The goal of squat is to provide extensions of common statistical methods for the
analysis of unit quaternion time series. Available statistical methods for QTS
samples are currently:

- random generation according to the Gaussian functional model via `rnorm_qts()`,
- `scale()`,
- `mean()`,
- `median()`,
- distance matrix computation via `distDTW()` (i.e. for now we use the dynamic time warping),
- tangent principal component analysis via `prcomp()`,
- k-means with optional alignment via `kmeans()`.

## Installation

You can install the package directly from CRAN:
```r
install.packages("squat")
```
or you can choose to install the development version from GitHub:
```r
# install.packages("remotes")
remotes::install_github("LMJL-Alea/squat")
```

The package has its own webpage: https://lmjl-alea.github.io/squat/.
