---
title: "rgudhi 0.1.0"
description: |
  An Interface to the GUDHI Library for Topological Data Analysis.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2023-01-23
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - topological data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="50%", fig.align='center'}
img <- png::readPNG("../../images/logo-gudhi.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

## Overview

[**rgudhi**](https://lmjl-alea.github.io/rgudhi/) `v0.1.0` provides an almost
full wrapper of the `v3.7.1` of the GUDHI library for topological data analysis.
Only the cover complex class is missing due to non-reproducibility issues with
random number generators. With GUDHI accessible from R,
[**rgudhi**](https://lmjl-alea.github.io/rgudhi/) `v0.1.0` features:

- data structure to encode simplicial complexes;
- computation of persistence diagrams;
- various usual preprocessing tools for persistence diagrams;
- a dedicated `S3` class [`persistence_diagram`](https://lmjl-alea.github.io/rgudhi/reference/persistence_diagram.html) for persistence diagram;
- [`plot()`](https://lmjl-alea.github.io/rgudhi/reference/plot.html) and [`ggplot2::autoplot()`](https://lmjl-alea.github.io/rgudhi/reference/plot.html) methods for [`persistence_diagram`](https://lmjl-alea.github.io/rgudhi/reference/persistence_diagram.html) objects;
- vector and kernel representations of persistence diagrams;
- a number of metrics to quantify distances between persistence diagrams 
([Bottleneck](https://lmjl-alea.github.io/rgudhi/reference/BottleneckDistance.html), [Persistence Fisher](https://lmjl-alea.github.io/rgudhi/reference/PersistenceFisherDistance.html), [Wasserstein](https://lmjl-alea.github.io/rgudhi/reference/WassersteinDistance.html), [Sliced Wasserstein](https://lmjl-alea.github.io/rgudhi/reference/SlicedWassersteinDistance.html)).
- functions to sample points from sphere ([`sphere()`](https://lmjl-alea.github.io/rgudhi/reference/sphere.html)) and torus ([`torus()`](https://lmjl-alea.github.io/rgudhi/reference/torus.html));
- a persistence-based clustering algorithm coined [Tomato](https://lmjl-alea.github.io/rgudhi/reference/Tomato.html).

The package also wraps all clustering algorithms from the [**sklearn.cluster**](https://scikit-learn.org/stable/modules/classes.html#module-sklearn.cluster)
module because they can be useful when using the `Atol` vectorization method for
persistence diagram.

It also wraps all scalers classes from
[**sklearn.preprocessing**](https://scikit-learn.org/stable/modules/classes.html#module-sklearn.preprocessing)
for use in various classes as well.

## Installation

You can install the package directly from CRAN:
```r
install.packages("rgudhi")
```
or you can choose to install the development version from GitHub:
```r
# install.packages("remotes")
remotes::install_github("LMJL-Alea/rgudhi")
```

The package has its own webpage: https://lmjl-alea.github.io/rgudhi/.
