---
title: "New paper on domain selection for functional data"
description: |
  Check out our newly published research which presents a unified framework to 
  perform domain selection for functional data with familywise error rate 
  control.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2022-04-27
output:
  distill::distill_article:
    self_contained: false
categories:
  - research
  - functional data
  - inference
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="80%", fig.align='center'}
img <- png::readPNG("../../images/biometrics_twt.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

Functional data are smooth, often continuous, random curves, which can be seen
as an extreme case of multivariate data with infinite dimensionality. Just as
component-wise inference for multivariate data naturally performs feature
selection, subset-wise inference for functional data performs domain selection.
In this paper, we present a unified testing framework for domain selection on
populations of functional data. In detail, $p$-values of hypothesis tests
performed on point-wise evaluations of functional data are suitably adjusted for
providing control of the family-wise error rate (FWER) over a family of subsets
of the domain. We show that several state-of-the-art domain selection methods
fit within this framework and differ from each other by the choice of the family
over which the control of the FWER is provided. In the existing literature,
these families are always defined a priori. In this work, we also propose a
novel approach, coined threshold-wise testing, in which the family of subsets is
instead built in a data-driven fashion. The method seamlessly generalizes to
multidimensional domains in contrast to methods based on a priori defined
families. We provide theoretical results with respect to consistency and control
of the FWER for the methods within the unified framework. We illustrate the
performance of the methods within the unified framework on simulated and real
data examples and compare their performance with other existing methods.

Find out more by reading our paper [@abramowicz2022domain] which can be found at <https://onlinelibrary.wiley.com/doi/10.1111/biom.13669>.
