---
title: "fdacluster 0.1.1"
description: |
  An R package for jointly aligning and clustering functional data
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2022-05-10
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - functional data
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="50%", fig.align='center'}
img <- png::readPNG("../../images/fdacluster_hex_trans.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

## Overview

The [**fdacluster**](https://astamm.github.io/fdacluster/index.html) package
currently provides the `kma()` function that jointly performs clustering and
alignment of a functional data set. The implementation follows [@sangalli2010k]
including improvements proposed in [@vantini2012definition].

The package is, in its current form, mainly a C++ re-implementation of the
[**fdakma**](https://cran.r-project.org/web/packages/fdakma/index.html) package
which makes it insanely faster. It will include more clustering methods in a
near future as well.

## Installation

You can install the released version of
[**fdacluster**](https://astamm.github.io/fdacluster/) from
[CRAN](https://cran.r-project.org) with:

```r
install.packages("fdacluster")
```

Alternatively you can install the development version of
[**fdacluster**](https://astamm.github.io/fdacluster/) from [GitHub](https://github.com/)
with:

```r
# install.packages("remotes")
remotes::install_github("astamm/fdacluster")
```
