---
title: "riot 1.0.0"
description: |
  An R interface for popular tractography file formats.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2022-05-10
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - connectome-valued functional data
  - data import
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="50%", fig.align='center'}
img <- png::readPNG("../../images/riot_hex_trans.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

## News

In this first major release, we:

* Added support to read
[MRtrix](https://mrtrix.readthedocs.io/en/latest/getting_started/image_data.html)
`.tck/.tsf` file formats.
* Added support to read [TrackVis](http://trackvis.org/docs/?subsect=fileformat)
`.trk` file formats.
* Use only one core to compile VTK for compliance with CRAN policy (thanks to
Prof. B. Ripley).
* Added tilde expansion on file paths.

We make it the first major release as we consider that the most popular
tractography formats are now supported by
[**riot**](https://astamm.github.io/riot/). We chose by design to support only
VTK and medInria file formats for writing.

## Installation

You can install the package directly from CRAN:
```r
install.packages("riot")
```
or you can choose to install the development version from GitHub:
```r
# install.packages("remotes")
remotes::install_github("astamm/riot")
```

The package has its own webpage: https://astamm.github.io/riot/.
