---
title: "nloptr 2.0.0"
description: |
  A new version of the nloptr package has been released! The NLopt library has 
  switched to CMake as build system. I am replacing Jelmer Ypma as maintainer of 
  the package.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2022-01-28
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - optimization
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="50%", fig.align='center'}
img <- png::readPNG("../../images/nloptr_hex_trans.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

> A new version of the [`nloptr`](https://astamm.github.io/nloptr/) package has
been released! The [NLopt](https://nlopt.readthedocs.io/en/latest/) library has
switched to [CMake](https://cmake.org) as build system. I am replacing Jelmer
Ypma as maintainer of the package.

## `nlopt`

The NLopt library [@nlopt] is a free/open-source library for nonlinear
optimization, providing a common interface for a number of different free
optimization routines available online as well as original implementations of
various other algorithms.

Starting from `v2.5.0`, NLopt has switched to `cmake` as build system, making
the latest versions incompatible with the way `nloptr` used to interface the
library.

## `nloptr`

As of `v2.0.0`, we now provide an R interface to `nlopt >= 2.7.0`. This is
achieved in different ways depending on the operating system you are using.

### Windows

There is no dependency required for Windows users.

- If you are running on `R <= 4.1.x`, installation of `nloptr` will rely on a
valid internet connexion to download the `nlopt 2.7.1` headers and static
library from `rwinlib` and use it.
- If you are running on `R >= 4.2.x`, installation of `nloptr` will directly
link with `nlopt 2.7.1` from the `Rtools42` toolchain.

### Linux and macOS

On Unix-like systems, installation of `nloptr` will go through the following
steps:

Is `pkg-config` available on your `PATH`?

- NO: compile `nlopt 2.7.1` from included sources using `cmake >= 3.15.0`.
- YES: can `pkg-config` detect a system build of `nlopt >= 2.7.0`?
  
  - NO: compile `nlopt 2.7.1` from included sources using `cmake >= 3.15.0`.
  - YES: install `nloptr` using the detected system build of `nlopt`.

Hence, `cmake >= 3.15.0` can become a system requirement for installing `nloptr`
on your machine if you are running on a Unix-like system without `pkg-config` or
without a system build of `nlopt >= 2.7.0`.
