---
title: "Mainstream Report #1"
description: |
  The first mainstream report for the MS-CSI project is now available online!
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 04-08-2021
output:
  distill::distill_article:
    self_contained: false
categories:
  - research
  - ms-csi
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(magrittr)
source("scripts/arrange_tree.R")
load("data/arrangeTree.Rdata")
```

```{r, fig.align='center', out.width="0%"}
dend <- arrange_tree(periores, ext_inf)
plot(dend)
```

The [MS-CSI](../../ms-csi.html) project is funded by the Association
pour la Recherche en Sclérose En Plaques (ARSEP) through a number of
French donors. It is important for the donors to have mainstream
material that they can read to understand what scientific discoveries we
were able to make thanks to their donations. A first mainstream report
for general audience is now available
[here](../../ms-csi/2021-04-08-msr1/index.html). It is intended to be
read by French native speakers and is thus written in French. In this
troubled period of global pandemic, the MS-CSI project has been delayed
and we expect it to start next September 2021. However, in the meantime,
a number of milestones have been achieved around our eGait solution that
we wanted to share with our generous donors and, more generally, to the
general audience who care about MS.
