---
title: "flipr 0.2.1"
description: |
  A new version of the flipr package has been released! Check it out if you are looking for a consistent, robust and easy way of doing permutation inference.
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 04-06-2021
output:
  distill::distill_article:
    self_contained: false
categories:
  - software
  - permutation inference
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r logo, out.width="50%", fig.align='center'}
img <- png::readPNG("../../images/flipr_hex_trans.png")
g <- grid::rasterGrob(img, interpolate = TRUE)
ggplot() +
  annotation_custom(g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  theme_void()
```

The permutation framework is particularly well suited for inferential
purposes as it allows one to do point estimation, confidence regions and
hypothesis tests. The [**flipr**](https://astamm.github.io/flipr/)
package makes it easy and fun to perform all these inferential tasks
within the permutation paradigm.

The central object is the so-called $p$-value function. The $p$-value
function for a set of parameters $\Theta$ is a curve that represents the
variation of the $p$-value of an hypothesis test in which the null
hypothesis is $\Theta = \Theta_0$ as a function of $\Theta_0$
[@martin2017; @fraser2019; @infanger2019].

Observe that, the non-parametric combination method available in the
permutation framework [@pesarin2010] allows you to infer multiple
parameters at once. This means that you can natively compute a single
confidence region for multiple parameters that controls the family-wise
error rate by construction (think of a confidence region for both the
mean and the variance for instance).

Version 0.2.1 of the [**flipr**](https://astamm.github.io/flipr/)
package released on
[CRAN](https://cran.r-project.org/web/packages/flipr/index.html)
features $4$ main functions:

-   You can compute $p$-value function when inferring multiple
    parameters at once via `two_sample_pf()`;
-   You can compute a single point estimate for a single parameter of
    interest via `two_sample_pe()`;
-   You can compute a confidence interval for a single parameter of
    interest via `two_sample_ci()`;
-   You can automatically draw the $p$-value function for a single
    parameter of interest via `two_sample_viz()`.

This version also implements a number of small changes and bug fixes,
the full list of which is available
[here](https://astamm.github.io/flipr/news/index.html).

Finally, a number of articles can be found on the dedicated website for
[**flipr**](https://astamm.github.io/flipr/):
<https://astamm.github.io/flipr/>. They explain in details how the
permutation framework can be used for statistical inference and how
[**flipr**](https://astamm.github.io/flipr/) offers easy tools for
making that happen seamlessly.
