---
title: "rdatagouv 0.1.0"
description: |
  The first version of the [**rdatagouv**](https://astamm.github.io/rdatagouv/) package
  has been released!
author:
  - name: Aymeric Stamm
    url: https://astamm.github.io
    affiliation: Department of Mathematics Jean Leray, UMR CNRS 6629
    affiliation_url: https://www.math.sciences.univ-nantes.fr
    orcid_id: 0000-0002-8725-3654
date: 2026-09-12
categories: [software, open data]
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = FALSE)
library(rdatagouv)
```

`rdatagouv` is an R client for the public API of
[data.gouv.fr](https://www.data.gouv.fr), the French government's open data
platform. It helps you *find* a dataset that matches your interests, *judge*
whether it is usable, *download* it, and later *re-fetch the exact same table*
reproducibly. Requests are built on top of the
[`httr2`](https://httr2.r-lib.org) package.

## Installation

The package is on CRAN under v0.1.0, so you can install it with:

```r
# install.packages("pak") # if you don't have pak yet
pak::pak("rdatagouv")
```

Alternatively, you can install the development version from GitHub:

```r
pak::pak("astamm/rdatagouv")
```

## Quick start

`rdatagouv` revolves around a simple workflow: **find** a dataset, **judge**
whether it is usable, **fetch** it into a table, and later **re-fetch the exact
same table** reproducibly. The examples below hit the live data.gouv.fr API and
show real results.

### 1. Find

As illustrated in @tbl-find-datasets, `dg_find_datasets()` searches the catalog by keyword and returns one row per
dataset, with metadata (`quality_score`, `views`, `license`, ...) to help you
choose:

```{r}
#| label: tbl-find-datasets
#| tbl-cap: "First 5 datasets matching the keyword *vélo* (bike) on data.gouv.fr"
#| echo: true
hits <- dg_find_datasets(q = "vélo", n = 5)
hits |> 
  dplyr::select(title, id, organization, quality_score) |>
  gt::gt() |> 
  gt::cols_label(
    title = "Title",
    id = "Dataset ID",
    organization = "Producer",
    quality_score = "Quality Score"
  ) |> 
  gt::fmt_percent(columns = "quality_score", decimals = 0) |>
  gt::opt_stylize(style = 4)
```

You can also discover producers with `dg_find_organization()` or curated themes
with `dg_find_topics()`, and narrow a search to one of them via the
`organization` / `topic` arguments.

### 2. Judge

Before pulling, glimpse a dataset's health and engagement in one call, or read
its documented columns with `dg_schema()`:

```{r}
#| label: glimpse
#| echo: true
g <- dg_glimpse("6a6be5976a05df136d48fb7a")
g$quality$score
```

### 3. Fetch

As illustrated in @tbl-fetch, `dg_pull_dataset()` downloads the dataset's first tabular resource and returns a
single tibble, with a stable table id attached as an attribute:

```{r}
#| label: tbl-fetch
#| tbl-cap: "First 5 rows of the C.vélo bike stations dataset"
#| echo: true
tbl <- dg_pull_dataset("6397c0ff56d3963118a18345")     # C.vélo bike stations
tbl |> 
  dplyr::select(station_id, name, physical_configuration, lat, lon, altitude) |>
  dplyr::slice_head(n = 5) |>
  gt::gt() |> 
  gt::cols_label(
    station_id = "Station ID",
    name = "Station Name",
    physical_configuration = "Physical Configuration",
    lat = "Latitude",
    lon = "Longitude",
    altitude = "Altitude"
  ) |> 
  gt::opt_stylize(style = 4)
```

At any time, you can retrieve the table's stable id with `dg_table_id()`:

```{r}
#| label: table-id
#| echo: true
dg_table_id(tbl)
```

### 4. Re-fetch reproducibly

Save that id, and `dg_refetch()` gets back the exact same table later — no
matter how titles or file names drift on the platform:

```{r}
#| label: refetch
#| echo: true
again <- dg_refetch(tbl)
identical(again, tbl)
```

### 5. Summarise

As illustrated in @tbl-summarise, `dg_summary()` computes metrics for one table (size, columns, rows,
missing-value rate), and `dg_summarise()` does the same for many at once — even
an entire `dg_find_datasets()` result:

```{r}
#| label: tbl-summarise
#| tbl-cap: "Summary of the `iris` and `mtcars` datasets"
#| echo: true
dg_summarise(datasets = list(iris = iris, mtcars = mtcars)) |> 
  gt::gt() |> 
  gt::cols_label(
    dataset = "Dataset",
    size_kb = "Size (KB)",
    n_vars = "Number of Variables",
    n_numeric = "Numeric Variables",
    n_non_numeric = "Non-numeric Variables",
    n_rows = "Number of Rows",
    prop_missing = "Missing Rate"
  ) |> 
  gt::fmt_number(columns = c("size_kb", "n_vars", "n_numeric", "n_non_numeric", "n_rows"), decimals = 0) |>
  gt::fmt_percent(columns = "prop_missing", decimals = 1) |>
  gt::opt_stylize(style = 4)
```

For a deeper tour — including column-type control and handling parsing
problems — see the [vignette](https://astamm.github.io/rdatagouv/articles/rdatagouv.html).

## Supported formats

`dg_pull_dataset()` downloads the first tabular resource of a dataset among
CSV, CSV.GZ, XLS, XLSX, PARQUET, TSV, TXT and JSON, and returns the parsed
table as a single tibble. A ZIP resource is unpacked and its first parseable
file is returned by default; `all_files = TRUE` keeps every contained file in
one of these formats as a named list — one element per file. The
delimiter of CSV/TXT resources is auto-detected (comma, semicolon, tab, pipe,
...), so both standard and European-style (semicolon/comma-decimal) files are
handled without special configuration. Column types are inferred by vroom, but
seeded by default from data.gouv's own `csv-detective` profile
(`use_tabular_types = TRUE`) and you can force specific columns with
`col_types = c(col = "Date")` (e.g. to make a mostly-padded ISO date column with
a few non-padded stragglers read as `Date`, turning the stragglers into `NA`;
explicit `col_types` always win). The profile is a best-effort signal — it only
exists for single-file resources indexed by the tabular service, and a missing
profile (or a ZIP member) falls back to type inference; pass
`use_tabular_types = FALSE` to disable seeding entirely. Any parsing issues are
attached to the returned table as an `rdatagouv_problems` attribute — read them
with `dg_problems(tbl)` — rather than a noisy per-cell warning.

The discovery catalog (`dg_find_datasets()`) is restricted to the official
tabular formats data.gouv.fr itself indexes (`csv`, `csv.gz`, `xls`, `xlsx`,
`parquet`), so every listed dataset is in principle openable as a table. Use
the `format` argument to narrow the catalog to datasets with a resource in a
specific format (the v2 search API matches several formats as a server-side
union). The exact `n_resources`/`formats`/`has_table`/`has_schema` resource
columns are populated by passing `resources = TRUE`. Direct pulls additionally
accept `tsv`, `txt` and `json` resources.

## Disclaimer

This package is not affiliated with or endorsed by data.gouv.fr. It is an
independent project, and the authors are not responsible for the content of
the datasets it indexes. This first version is a proof of concept and may
contain bugs. Please report any issues on the
[GitHub repository](https://github.com/astamm/rdatagouv/issues).

This package is the result of a joint effort initiated during the French
[Finist'R](https://stateofther.pages-forge.inrae.fr/finistr2026/) bootcamp which
was held in Roscoff in August 2026. It served as a complex task to benchmark R
package development assisted by AI. The refinements that came up during the
bootcamp were made with the assistance of `deepseek-v4-flash` as provided by
Albert API, the French government's AI platform. The author is grateful to the
Finist'R organizers and the Albert API team for their support and guidance.